---
title: "LADAL Concordancing Tool"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: hide
    df_print: paged
---

```{css, echo=FALSE}
/* ── LADAL brand colours ───────────────────────────────────────────── */
:root {
  --ladal-purple:  #51247a;
  --ladal-light:   #f4f0f8;
  --ladal-accent:  #8e44ad;
  --ladal-gold:    #f0a500;
  --ladal-success: #27ae60;
  --ladal-info:    #d6eaf8;
  --ladal-border:  #d7d1cc;
}

/* Page header */
h1.title {
  color: var(--ladal-purple);
  border-bottom: 4px solid var(--ladal-gold);
  padding-bottom: 0.4em;
}
h2, h3 { color: var(--ladal-purple); }

/* Info / warning boxes */
.info-box {
  background-color: var(--ladal-light);
  border-left: 5px solid var(--ladal-purple);
  border-radius: 4px;
  padding: 0.9em 1.2em;
  margin: 1em 0;
  color: #333;
}
.info-box strong { color: var(--ladal-purple); }

.success-box {
  background-color: #eafaf1;
  border-left: 5px solid var(--ladal-success);
  border-radius: 4px;
  padding: 0.9em 1.2em;
  margin: 1em 0;
}

.tip-box {
  background-color: #fef9e7;
  border-left: 5px solid var(--ladal-gold);
  border-radius: 4px;
  padding: 0.9em 1.2em;
  margin: 1em 0;
}

/* Parameter table */
.param-table th {
  background-color: var(--ladal-purple);
  color: white;
}

/* Keyword highlight inside kwic output */
.keyword { color: var(--ladal-accent); font-weight: bold; }

/* Keep code chunks readable */
pre { background-color: #f8f6fb !important; }
```

---

<div class="info-box">
📖 **About this tool** — This notebook accompanies the
[LADAL tutorial *Concordancing with R*](https://ladal.edu.au/kwics.html).
It lets you upload your own plain-text files and run keyword-in-context
(KWIC) concordance searches, then explore and export your results.
All code is **folded by default** — click **Code** on the right of any
chunk to expand it.
</div>

---

## 1 · Setup

```{r setup, message=FALSE, warning=FALSE}
# ── Packages ──────────────────────────────────────────────────────────
required_pkgs <- c("quanteda", "tidyverse", "writexl", "here", "DT", "knitr")

for (pkg in required_pkgs) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

library(quanteda)
library(tidyverse)
library(writexl)
library(here)
library(DT)
library(knitr)

# Silence quanteda startup messages
quanteda_options(verbose = FALSE)
```

---

## 2 · Load your texts

<div class="info-box">
📂 **How to upload your own texts**

1. Open the **`MyTexts`** folder in the file browser on the left.
2. **Drag and drop** your `.txt` files into that folder.
3. Run the code chunk below — it will read every `.txt` file it finds.

> Only plain-text (`.txt`) files are supported. One file = one "document"
> in the concordance. File names become document IDs.
</div>

```{r load-texts, message=FALSE, warning=FALSE}
# ── Helper: load all .txt files from a folder ─────────────────────────
load_texts <- function(folder = "notebooks/MyTexts") {
  txt_files <- list.files(folder, pattern = "\\.txt$",
                          full.names = TRUE, ignore.case = TRUE)

  if (length(txt_files) == 0) {
    message("⚠  No .txt files found in '", folder, "'.")
    message("   Falling back to a built-in demo corpus (quanteda::data_corpus_inaugural).")
    corp <- corpus(quanteda::data_corpus_inaugural)
    return(corp)
  }

  raw <- map(txt_files, ~ readLines(.x, warn = FALSE) |> paste(collapse = " "))
  names(raw) <- tools::file_path_sans_ext(basename(txt_files))
  corp <- corpus(unlist(raw), docnames = names(raw))
  message("✔  Loaded ", length(txt_files), " file(s): ",
          paste(names(raw), collapse = ", "))
  corp
}

# ── Load ──────────────────────────────────────────────────────────────
corp <- load_texts("notebooks/MyTexts")

# Quick summary
summary(corp, n = 10) |>
  select(Text, Types, Tokens, Sentences) |>
  kable(caption = "Loaded documents (first 10 shown)")
```

---

## 3 · Run a concordance search

<div class="tip-box">
⚙️ **Customise your search** — Edit the parameters in the chunk below:

| Parameter | What it does | Example |
|---|---|---|
| `search_pattern` | The word / phrase to find | `"climate"`, `"the economy"` |
| `context_window` | Words shown left & right of keyword | `5`, `10` |
| `match_type` | How the pattern is interpreted | `"regex"`, `"glob"`, `"fixed"` |
| `ignore_case` | Ignore upper/lower case | `TRUE` / `FALSE` |
| `max_display` | Rows shown in the table | `50`, `100`, `Inf` |
</div>

```{r concordance, message=FALSE, warning=FALSE}
# ══════════════════════════════════════════════════════════════════════
#  ▶  EDIT THESE PARAMETERS TO CUSTOMISE YOUR SEARCH
# ══════════════════════════════════════════════════════════════════════

search_pattern  <- "the"   # word, phrase, or regex pattern
context_window  <- 5       # words left and right
match_type      <- "regex" # "regex" | "glob" | "fixed"
ignore_case     <- TRUE    # TRUE = case insensitive
max_display     <- 100     # max rows shown in interactive table

# ══════════════════════════════════════════════════════════════════════
#  (no need to edit below this line)
# ══════════════════════════════════════════════════════════════════════

mykwic <- kwic(
  tokens(corp),
  pattern        = phrase(search_pattern),
  window         = context_window,
  valuetype      = match_type,
  separator      = " ",
  case_insensitive = ignore_case
)

if (nrow(mykwic) == 0) {
  message("⚠  No matches found for pattern: '", search_pattern, "'")
} else {
  message("✔  Found ", nrow(mykwic), " match(es) for '", search_pattern, "'")
}
```

---

## 4 · Frequency summary

```{r frequency-summary, message=FALSE, warning=FALSE}
if (nrow(mykwic) > 0) {
  freq_summary <- as.data.frame(mykwic) |>
    group_by(docname) |>
    summarise(
      Hits      = n(),
      .groups   = "drop"
    ) |>
    arrange(desc(Hits)) |>
    rename(Document = docname)

  total_row <- tibble(Document = "**TOTAL**", Hits = sum(freq_summary$Hits))
  freq_table <- bind_rows(freq_summary, total_row)

  kable(freq_table,
        caption = paste0("Hit counts per document — pattern: '",
                         search_pattern, "'"))
} else {
  message("No results to summarise.")
}
```

---

## 5 · Browse results

<div class="info-box">
🔍 **Interactive table** — You can:

- **Sort** any column by clicking its header (▲ / ▼)
- **Filter by document** using the search box above each column
- **Search** across all columns using the global search box (top right)
- Use **Previous / Next** to page through results
</div>

```{r results-table, message=FALSE, warning=FALSE}
if (nrow(mykwic) > 0) {
  kwic_df <- as.data.frame(mykwic) |>
    select(
      Document  = docname,
      Position  = from,
      Left      = pre,
      Keyword   = keyword,
      Right     = post,
      Pattern   = pattern
    ) |>
    slice_head(n = max_display)

  datatable(
    kwic_df,
    rownames   = FALSE,
    filter     = "top",          # per-column filter boxes
    extensions = c("Buttons", "Scroller"),
    options    = list(
      dom          = "Bfrtip",
      buttons      = list("copy", "csv", "excel"),
      scrollY      = 420,
      scroller     = TRUE,
      pageLength   = 25,
      autoWidth    = TRUE,
      columnDefs   = list(
        list(className = "dt-left",  targets = c(2, 3, 4)),   # L/Kw/R left-aligned
        list(className = "dt-center", targets = c(0, 1, 5))   # doc/pos/pat centred
      )
    ),
    caption = htmltools::tags$caption(
      style = "color:#51247a; font-weight:bold;",
      paste0("KWIC results for '", search_pattern,
             "' (showing up to ", max_display, " of ",
             nrow(mykwic), " hits)")
    )
  ) |>
  formatStyle(
    "Keyword",
    color      = "#8e44ad",
    fontWeight = "bold"
  )
} else {
  cat("No results to display.")
}
```

---

## 6 · Export results

<div class="success-box">
💾 **Saving your concordance** — Run the chunk below to write an Excel
file to `notebooks/MyOutput/mykwic.xlsx`.  
Then right-click the file in the left-hand file browser and choose
**Download**.
</div>

```{r export, message=FALSE, warning=FALSE}
if (nrow(mykwic) > 0) {

  # ── Build a clean export table ─────────────────────────────────────
  export_df <- as.data.frame(mykwic) |>
    select(
      Document = docname,
      Position = from,
      Left     = pre,
      Keyword  = keyword,
      Right    = post,
      Pattern  = pattern
    )

  # ── Write Excel file ───────────────────────────────────────────────
  out_dir  <- here::here("notebooks/MyOutput")
  dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
  out_path <- file.path(out_dir, "mykwic.xlsx")

  write_xlsx(export_df, out_path)

  message("✔  Saved ", nrow(export_df), " rows to:\n   ", out_path)

} else {
  message("⚠  Nothing to export — no matches were found.")
}
```

<div class="success-box">
Also written as a CSV (useful if you don't have Excel):
</div>

```{r export-csv, message=FALSE, warning=FALSE}
if (nrow(mykwic) > 0) {
  csv_path <- file.path(here::here("notebooks/MyOutput"), "mykwic.csv")
  write_csv(export_df, csv_path)
  message("✔  CSV saved to: ", csv_path)
}
```

---

## 7 · Advanced: multiple searches

<div class="tip-box">
🔁 **Run several searches at once** — add as many patterns as you like
to `patterns_list` and re-run the chunk. All results are stacked into
one table and exported together.
</div>

```{r multi-search, message=FALSE, warning=FALSE}
# ══════════════════════════════════════════════════════════════════════
#  ▶  EDIT: list of search patterns
# ══════════════════════════════════════════════════════════════════════
patterns_list <- c("the", "a")

# ══════════════════════════════════════════════════════════════════════

multi_kwic <- map_dfr(patterns_list, function(pat) {
  hits <- kwic(
    tokens(corp),
    pattern          = phrase(pat),
    window           = context_window,
    valuetype        = match_type,
    case_insensitive = ignore_case
  ) |> as.data.frame()

  if (nrow(hits) > 0) hits else NULL
})

if (!is.null(multi_kwic) && nrow(multi_kwic) > 0) {
  cat("Multi-pattern hit counts:\n")
  multi_kwic |>
    count(pattern, name = "Hits") |>
    arrange(desc(Hits)) |>
    kable()
} else {
  cat("No hits across any of the supplied patterns.")
}
```

---

## Citation & Session Info

```{r citation, echo=FALSE}
cat(
'@manual{schweinberger2024kwictool,
  author       = {Schweinberger, Martin},
  title        = {LADAL Concordancing Tool},
  note         = {https://ladal.edu.au/tools.html},
  year         = {2024},
  organization = {The University of Queensland, School of Languages and Cultures},
  address      = {Brisbane},
  edition      = {2024.04.21}
}')
```

> Schweinberger, Martin. (2024). *LADAL Concordancing Tool*. Brisbane:
> The University of Queensland.
> <https://ladal.edu.au/tools.html> (Version 2024.04.21).

---

[LADAL Concordancing tutorial](https://ladal.edu.au/kwics.html) ·
[LADAL home](https://ladal.edu.au)

```{r session-info, collapse=TRUE}
sessionInfo()
```
